Using spatially biased variables in ecosystem condition accounting with a GIS based workflow

### 2024-01-29 manuscript/HIAs.qmd
explicit eval statements

<!--# The document is set to eval=F in the yaml, and some code chunks are set manually to eval=T. When working on the manuscript, you can run all (Ctr+Alt+R). This will read inn data cashed locally. To also evaluate the more time consuming chunks, change the YAML to eval: true and run all (or render to html):  -->

@@

```{r getRelevantNTs}
# I already did some work to identify the relevant nature types

@@

```{r relevantNTs}
# These are the nature types I'm keeping for 7FA.
# Note the absence of ombrotriphic types
alienNTS <- c("Slåttemyr",
              "Semi-naturlig våteng",
              "Rik åpen jordvannsmyr i mellomboreal sone",
              "Rik åpen sørlig jordvannsmyr",
              "Rik åpen jordvannsmyr i nordboreal og lavalpin sone")

@@

# and also to get the data coverage polygon. 
# These I had to manually download from naturbase.no 
dk <-   read_sf("../data/dekningskartNF/Naturtyper NiN - Dekningskart.shp") |>
  rbind(read_sf("../data/dekningskartTR/Naturtyper NiN - Dekningskart.shp")) |>
  rbind(read_sf("../data/dekningskartME/Naturtyper NiN - Dekningskart.shp")) |> 
  st_transform(st_crs(muni3)) |>
  st_intersection(muni3) |>
  st_union()

@@

```{r positionMap}

@@

```{r mireDataTestCrop}

@@

```{r mireTerra}

@@

```{r dk2-CoverageMaps}
# calculate area of survey coverage maps
# values goes into summary table in the ms
dk2 <-   read_sf("../data/dekningskartNF/Naturtyper NiN - Dekningskart.shp") |>
  rbind(read_sf("../data/dekningskartTR/Naturtyper NiN - Dekningskart.shp")) |>
  rbind(read_sf("../data/dekningskartME/Naturtyper NiN - Dekningskart.shp")) |>
  st_transform(st_crs(muni3)) |>
  st_intersection(muni3) |>
  st_cast("POLYGON") |>
  group_by(Municipality) |>
  summarise(geometry = st_union(geometry)) |>
  mutate(dk_area_km = geometry |> st_area(),
         dk_area_km = round(units::drop_units(dk_area_km * 1e-6)))
```

```{r mireArea}
# calculate area of the mires in each municipality
# -- Nordre Follo
mireArea <- mire_nf_terra |>
  global(c("mean", "sum"), na.rm=T) |>
  add_column("Municipality" = "Nordre Follo") |>
  mutate(mirePercent = round(mean*100, 1),
         mire_km2 = sum/1e+4)

@@

```{r infrastructureIndex}
# this data is on a 100x100m grid
infra <- stars::read_stars(path_infrastructure) |>
  # , which is mor then we need - warp it to 1x1km
  st_warp(cellsize = c(1000, 1000), 
                                   crs = st_crs(nf), 
                                   use_gdal = TRUE,
                                   method = "average") |>
  setNames("infrastructureIndex") |>
  st_transform(myCRS) |>
   mutate(infrastructureIndex = case_when(
    infrastructureIndex < 1 ~ 0,
    infrastructureIndex < 6 ~ 1,
    infrastructureIndex < 12 ~ 2,
    infrastructureIndex >= 12 ~ 3)) |>
  # takeing away point in the sea
  st_crop(outline)

@@

```{r readInfraStructureIndex}

@@

```{r infraMuniMaps}
infraMuniMap <- tm_shape(muni3)+
  tm_borders()+
tm_shape(infraMuni3)+
  tm_polygons(col= "infrastructureIndex",
              style="cat",
              title = "Human Impact\nArea Class")+
  tm_shape(muni3)+
  tm_borders(lwd=3, col="black")+
  tm_facets(by = "Municipality")+
  tm_layout(legend.show=TRUE)

@@

```{r muni-tbl}

>>>>

<!--# The document is set to eval=F in the yaml, and some code chunks are set manually to eval=T. When working on the manuscript, you cannot run all with  Ctr+Alt+R as this ignores the YAML. This will read inn data cashed locally. To also evaluate the more time consuming chunks, change the YAML to eval: true and run all (or render to html):  -->

@@

```{r getRelevantNTs}
#| eval: false

# I already did some work to identify the relevant nature types

@@

```{r relevantNTs}
#| eval: false

# These are the nature types I'm keeping for 7FA.
# Note the absence of ombrotriphic types
alienNTS <- c("Slåttemyr",
              "Semi-naturlig våteng",
              "Rik åpen jordvannsmyr i mellomboreal sone",
              "Rik åpen sørlig jordvannsmyr",
              "Rik åpen jordvannsmyr i nordboreal og lavalpin sone")

@@

# and also to get the data coverage polygon. 
# These I had to manually download from naturbase.no 
dk <-   read_sf("../data/dekningskartNF/Naturtyper NiN - Dekningskart.shp") |>
  rbind(read_sf("../data/dekningskartTR/Naturtyper NiN - Dekningskart.shp")) |>
  rbind(read_sf("../data/dekningskartME/Naturtyper NiN - Dekningskart.shp")) |> 
  st_transform(st_crs(muni3)) |>
  st_intersection(muni3) |>
  st_union()
```

@@

```{r positionMap}
#| eval: false

@@

```{r mireDataTestCrop}
#| eval: false

@@

```{r mireTerra}
#| eval: false

@@

```{r dk2-CoverageMaps}
#| eval: false

# calculate area of survey coverage maps
# values goes into summary table in the ms
dk2 <-   read_sf("../data/dekningskartNF/Naturtyper NiN - Dekningskart.shp") |>
  rbind(read_sf("../data/dekningskartTR/Naturtyper NiN - Dekningskart.shp")) |>
  rbind(read_sf("../data/dekningskartME/Naturtyper NiN - Dekningskart.shp")) |>
  st_transform(st_crs(muni3)) |>
  st_intersection(muni3) |>
  st_cast("POLYGON") |>
  group_by(Municipality) |>
  summarise(geometry = st_union(geometry)) |>
  mutate(dk_area_km = geometry |> st_area(),
         dk_area_km = round(units::drop_units(dk_area_km * 1e-6)))
```

```{r mireArea}
#| eval: false

# calculate area of the mires in each municipality
# -- Nordre Follo
mireArea <- mire_nf_terra |>
  global(c("mean", "sum"), na.rm=T) |>
  add_column("Municipality" = "Nordre Follo") |>
  mutate(mirePercent = round(mean*100, 1),
         mire_km2 = sum/1e+4)

@@

```{r infrastructureIndex}
#| eval: false

# this data is on a 100x100m grid
infra <- stars::read_stars(path_infrastructure) |>
  # , which is mor then we need - warp it to 1x1km
  st_warp(cellsize = c(1000, 1000), 
                                   crs = st_crs(nf), 
                                   use_gdal = TRUE,
                                   method = "average") |>
  setNames("infrastructureIndex") |>
  st_transform(myCRS) |>
   mutate(infrastructureIndex = case_when(
    infrastructureIndex < 1 ~ 0,
    infrastructureIndex < 6 ~ 1,
    infrastructureIndex < 12 ~ 2,
    infrastructureIndex >= 12 ~ 3)) |>
  # takeing away point in the sea
  st_crop(outline)

@@

```{r readInfraStructureIndex}
#| eval: false

@@

```{r infraMuniMaps}
#| eval: false

infraMuniMap <- tm_shape(muni3)+
  tm_borders()+
tm_shape(infraMuni3)+
  tm_polygons(col= "infrastructureIndex",
              style="cat",
              title = "Human Impact\nArea Class")+
  tm_shape(muni3)+
  tm_borders(lwd=3, col="black")+
  tm_facets(by = "Municipality")+
  tm_layout(legend.show=TRUE)

@@

<!--# add mean_HIA to the mix -->

```{r muni-tbl}
#| eval: false
